Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-12 paper/paper.tex
Typos

Additionally,
the DMN was proposed to be situated
at the top of the brain network hierarchy, with
the subordinate salience and dorsal attention network in the middle and
the primary sensory cortices at the bottom
\citep{carhart2010default, margulies2016situating}.
Its putative involvement in thinking about 
hypothetical experiences and future outcomes
appears to tie in with the implicit computation of
action and state cascades as a function of
experienced events and collected feedback from the past.
A policy matrix encapsulates the choice probabilities of possible actions
on the world given a current situation (i.e., state).
The DMN may subserve
  constant exploration of candidate action trajectories and
  nested estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides a potential explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.

@@

We therefore hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN. First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error. Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association. We further expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.

Our process model of the DMN has also implications for experiments in neuroeconomy; especially for temporal discounting and continuous learning paradigms. More specifically, we hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry. During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a multi-step learning paradigm. We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment. In the case of temporal discounting, we hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
\citep{luksys2009stress, haushofer2014psychology}.
Using such a delayed-reward experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced back to its corresponding neural representation. Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age, education, IQ, and personality traits.
% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).

@@

A last experimental prediction for future studies concerns the role of synaptic epigenesis in the creation of new entries in policy matrix. Indeed, we did not adress here the additional layer of learning which concerns the addition of new entries in the state and action spaces. Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}. The tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}. Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}. To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synaptic changes compare to other parts of the brain.

>>>>

Additionally,
the DMN was proposed to be situated
at the top of the brain network hierarchy, with
the subordinate salience and dorsal attention network in the middle and
the primary sensory cortices at the bottom
\citep{carhart2010default, margulies2016situating}.
Its putative involvement in thinking about 
hypothetical experiences and future outcomes
appears to tie in with the implicit computation of
action and state cascades as a function of
experienced events and collected feedback from the past.
A policy matrix encapsulates the choice probabilities of possible actions
on the world given a current situation (i.e., state).
The DMN may subserve
  constant exploration of candidate action trajectories and
  nested estimation of their
  cumulative reward outcomes. Implicit computation of future choices
  provides a potential explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.

@@

We therefore hypothesize in humans a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN. First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error. Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association. We further expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.

Our process model of the DMN has also implications for experiments in neuroeconomy; especially for temporal discounting and continuous learning paradigms. More specifically, we hypothesize in humans a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry. During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a multi-step learning paradigm. We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment. In the case of temporal discounting, we hypothesize in humans that the relevant time horizon is modulated by various factors such as age, acute stress, and time-enduring impulsivity traits
\citep{luksys2009stress, haushofer2014psychology}.
Using such a delayed-reward experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced back to its corresponding neural representation. Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age, education, IQ, and personality traits.
% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).

@@

A last experimental prediction for future studies concerns how synaptic epigenesis may shape the policy matrix. Indeed, we did not adress here the additional layer of learning which concerns the addition of new entries in the state and action spaces. Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}. The tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}. Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}. To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synaptic changes compare to other parts of the brain.
